Draft Exploration

### 2020-04-19 draft_exploration.Rmd
war and career length table and boxplots

war <- batting_war %>%
  full_join(pitching_war, by = c("mlb_ID", "year_ID")) %>%
  mutate(war = bat_war + pitch_war)
```

@@

mlb_career <- mlb_start %>%
  left_join(mlb_end, by = "mlb_ID") %>%
  mutate(mlb_length = mlb_end - mlb_start + 1) %>%
  select(mlb_ID, mlb_length)

@@

debut_age(max_round = 20, startyear = 1990, endyear = 2011, min_round = 1)

@@

>>>>

war <- batting_war %>%
  full_join(pitching_war, by = c("mlb_ID", "year_ID")) %>%
  mutate(bat_war = ifelse(is.na(bat_war), 0, bat_war)) %>%
  mutate(pitch_war = ifelse(is.na(pitch_war), 0, pitch_war)) %>%
  mutate(war = bat_war + pitch_war)
```

@@

mlb_career <- mlb_start %>%
  left_join(mlb_end, by = "mlb_ID") %>%
  mutate(mlb_length = mlb_end - mlb_start + 1) %>%
  select(mlb_ID, mlb_end, mlb_length)

@@

debut_age(max_round = 40, startyear = 1990, endyear = 2011, min_round = 1)

@@

# I created hs_mlb_length which takes all high school picks who reached majors
# and have not been in the majors since 2017 latest to make it very likely their
# MLB career is over. I then grouped by whether they signed out of high school
# and found some summary stats for the career lengths of each group and career
# WAR.

hs_mlb_length <- master %>%
  filter(source == "H" & draft_numb == 1 & high_level == "MLB") %>%
  filter(mlb_end <= 2017) %>%
  group_by(signed) %>%
  summarize(q1 = quantile(mlb_length, 0.25),
            median_length = median(mlb_length),
            mean_length = mean(mlb_length),
            q3 = quantile(mlb_length, 0.75))

hs_mlb_war <- master %>%
  filter(source == "H" & draft_numb == 1 & high_level == "MLB") %>%
  filter(mlb_end <= 2017) %>%
  group_by(signed) %>%
  summarize(q1 = quantile(war, 0.25),
            median_war = median(war),
            mean_war = mean(war),
            q3 = quantile(war, 0.75))

# Made a gt table of the summary data from the above hs_mlb_war table

gt(data = hs_mlb_war) %>%
  tab_header(
    title = "Career Wins Above Replacement of High School Picks to Reach MLB"
  ) %>%
  tab_spanner(
    columns = vars(signed, q1, median_war, mean_war, q3),
    label = "Grouped by whether player signed out of high school"
  ) %>%
  cols_label(
    signed = "Signed?",
    q1 = "25th Percentile",
    median_war = "Median",
    mean_war = "Mean",
    q3 = "75th Percentile"
  ) %>%
  fmt_number(
    columns = vars(q1, median_war, mean_war, q3),
    decimals = 2
  )

hs_mlb_inactive <- master %>%
  filter(source == "H" & draft_numb == 1 & high_level == "MLB") %>%
  filter(mlb_end <= 2017)

ggplot(hs_mlb_inactive, aes(signed, mlb_length, fill = signed)) +
  geom_boxplot() +
  coord_flip() +
  theme_light() +
  labs(
    x = "Years Played in MLB",
    y = "Signed out of High School?",
    title = "MLB Career Length Summary of High School Picks to Reach MLB",
    subtitle = "Grouped by whether player signed out of high school"
  )

ggplot(hs_mlb_inactive, aes(signed, war/mlb_length, fill = signed)) +
  geom_boxplot() +
  coord_flip() +
  theme_light() +
  labs(
    x = "Career MLB Wins Above Replacement",
    y = "Signed out of High School?",
    title = "MLB Career Wins Above Replacement of High School Picks to Reach MLB",
    subtitle = "Grouped by whether player signed out of high school"
  )
